---
title: "Google Analytics report for Very Statisticious"
author: "Ariel Muldoon"
date: "`r Sys.Date()`"
output: pdf_document
params:
     startdate: "2020-06-05"
     enddate: "2020-06-11"
---

```{r setup, include = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

library(googleAnalyticsR)
library(dplyr) # data manip
library(tidyr) # joins
library(ggplot2) # plotting
library(sf) # plotting maps
library(forcats) # factors

# Authorization
ga_auth(email = "ariel.muldoon@gmail.com")
```

I started tracking my blog Very Statisticious, aosmith.rbind.io, with Google Analytics on 2020-06-04.  Now I can track users and page views and look at how many and where folks are using the blog.

## Google analytics data

Pull Google Analytics data from web and analyze key metrics of interest.  Dates can be changed using the parameters.

I'm collecting number of unique page views, users and sessions by country, and users and sessions by state/city (US). For location data I remove users without data.

```{r data, echo = FALSE, message = FALSE, warning = FALSE}
vs_id = 220230987
pageviews = google_analytics(vs_id,
                             date_range = c(params$startdate, params$enddate),
                             metrics = "uniquePageViews",
                             dimensions = "pagePath")

# Country
# Remove those without location info
country = google_analytics(vs_id,
                           date_range = c(params$startdate, params$enddate),
                           metrics = c("users", "sessions"),
                           dimensions = "country") %>%
     filter(country != "(not set)")

# US States
# Remove those without location info
us_states = google_analytics(vs_id,
                             date_range = c(params$startdate, params$enddate),
                             metrics = c("users", "sessions"),
                             dimensions = c("country", "region"),
                             dim_filters = filter_clause_ga4(list(dim_filter("country", "EXACT", "United States", not = FALSE) ) ) ) %>%
     filter(region != "(not set)")

# US cities
# Remove those without location info
us_cities = google_analytics(vs_id,
                             date_range = c(params$startdate, params$enddate),
                             metrics = c("users", "sessions"),
                             dimensions = c("country", "region", "city"),
                             dim_filters = filter_clause_ga4(list(dim_filter("country", "EXACT", "United States", not = FALSE) ) ),
                             max = -1) %>%
     filter(country != "(not set)")

```

## Data cleaning

Clicks from facebook, in particular, changes the page ID.  These need to be cleaned.  In addition, I want to add "tag" info for pages that involve **ggplot2**, **emmeans**, or **simulations**.

```{r clean, echo = FALSE, warning = FALSE, message = FALSE}
cleanpage = pageviews %>%
     mutate(pagePath = gsub("\\?.*", "", pagePath) ) %>%
     group_by(pagePath) %>%
     summarise(uniquePageViews = sum(uniquePageViews) ) %>%
     filter(pagePath != "/", !grepl("tags|page", pagePath) )


# Create look-up table with page and tag 
     # (ggplot2, simulation, emmeans only, any more "other")
tag = tribble(~pagePath, ~tag,
              "/2018/01/09/simulate-simulate-part1/", "simulation",
              "/2018/01/09/simulate-simulate-part2/", "simulation",
              "/2018/01/19/reversing-the-order-of-a-ggplot2-legend/", "ggplot2",
              "/2018/04/23/simulate-simulate-part-2/", "simulation",
              "/2018/06/05/a-closer-look-at-replicate-and-purrr/", "simulation",
              "/2018/07/18/simulate-poisson-edition/", "simulation",
              "/2018/07/19/manual-legends-ggplot2/", "ggplot2",
              "/2018/08/20/automating-exploratory-plots/", "ggplot2",
              "/2018/08/29/getting-started-simulating-data/", "simulation",
              "/2018/11/16/plot-fitted-lines/", "ggplot2",
              "/2019/03/25/getting-started-with-emmeans/", "emmeans",
              "/2019/04/15/custom-contrasts-emmeans/", "emmeans",
              "/2019/04/22/embedding-subplots/", "ggplot2",
              "/2019/05/13/small-multiples-plot/", "ggplot2",
              "/2019/10/14/background-color_gradient/", "ggplot2"
)

# Add tag info to pagepath
     # Make any without tag "other"
pagetag = left_join(cleanpage, tag, by = "pagePath") %>%
     mutate(tag = fct_explicit_na(tag, na_level = "other") ) %>%
     rename(views = uniquePageViews)
```

Here's the cleaned table.

```{r cleantab}
knitr::kable(pagetag)
```

## Page visit stats

Here are page visit stats for `r params$startdate` through `r params$enddate`.

### Proportion of times the page was visited.

```{r proppage, echo = FALSE}
pagetag %>%
     arrange( desc(views) ) %>%
     mutate(prop = views/sum(views) ) %>%
     knitr::kable()
```

### Proportion of time each tag visited.

```{r proptag, echo = FALSE}
pagetag %>%
     group_by(tag) %>%
     summarise(total = sum(views) ) %>%
     arrange( desc(total) ) %>%
     mutate(prop = total/sum(total) ) %>%
     knitr::kable()
```

### Weighted proportions for tags, since some tags have more posts than others.

```{r proptag2, echo = FALSE}
pagetag %>%
     group_by(tag) %>%
     summarise(total = sum(views),
               n = n() ) %>%
     arrange( desc(total) ) %>%
     mutate(prop = (total/n)/(sum(total/n) ) ) %>%
     knitr::kable()
```

## Plot of use across world

```{r world, echo = FALSE, fig.height = 6, fig.width = 10}
# World map
world_sf = st_as_sf(rworldmap::getMap(resolution = "low"))

# This world map uses some diff country names
     # than what I have
# I believe Svalbard & Jan Mayen is a part of Norway
# Other than that world = ga_country 
     # United States of America = United States,
     # Republic of Serbia = Serbia,
     # Czech Republic = Czechia
     # Myanmar = Myanmar (Burma)
     # United Republic of Tanzania = Tanzania

country = country %>%
     mutate(ADMIN = case_when(country == "United States" ~ "United States of America",
                           country == "Serbia" ~ "Republic of Serbia",
                           country == "Myanmar (Burma)" ~ "Myanmar",
                           country == "Czechia" ~ "Czech Republic",
                           country == "Tanzania" ~ "United Republic of Tanzania",
                           country == "Svalbard & Jan Mayen" ~ "Norway",
                           TRUE ~ country) ) %>%
     group_by(ADMIN) %>%
     summarise(across(.cols = users:sessions, sum) )

# Add in user info to world map data
world_users = left_join(world_sf, country)

# Try Robinson projection based on Wilke gist
     # https://gist.github.com/clauswilke/783e1a8ee3233775c9c3b8bfe531e28a
crs_robin = "+proj=robin +lat_0=0 +lon_0=0 +x0=0 +y0=0"
# projection outline in long-lat coordinates
lats = c(90:-90, -90:90, 90)
longs = c(rep(c(180, -180), each = 181), 180)
robin_outline = 
     list(cbind(longs, lats)) %>%
     st_polygon() %>%
     st_sfc(
          crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"
     ) %>% 
     st_transform(crs = crs_robin)
# bounding box in transformed coordinates
xlim = c(-18494733, 18613795)
ylim = c(-9473396, 9188587)
robin_bbox = 
     list(
          cbind(
               c(xlim[1], xlim[2], xlim[2], xlim[1], xlim[1]), 
               c(ylim[1], ylim[1], ylim[2], ylim[2], ylim[1])
          )
     ) %>%
     st_polygon() %>%
     st_sfc(crs = crs_robin)
# area outside the earth outline
robin_without = st_difference(robin_bbox, robin_outline)

# Different breaks if 7 days or 30 days
     # (best guess for now)
if(diff(c(as.Date(params$startdate), as.Date(params$enddata))) > 7) {
     leg_breaks = c(10, 100, 1000, 10000) 
     } else {
     leg_breaks = c(10, 100, 500)
     }
# Sequential color scheme example: https://wilkelab.org/practicalgg/articles/Texas_income.html
world_plot = ggplot() + 
     geom_sf(data = robin_without, fill = "white", color = NA) +
     geom_sf(data = robin_outline, fill = NA, color = "black", size = 0.5/.pt) +
     geom_sf(data = world_users,
             aes(fill = users), color = NA) +
     geom_sf(data = world_users, fill = NA, color = "black", size = 0.5/.pt) +
     scale_x_continuous(name = NULL, breaks = seq(-120, 120, by = 60)) +
     scale_y_continuous(name = NULL, breaks = seq(-60, 60, by = 30)) +
     coord_sf(xlim = 0.95*xlim, ylim = 0.95*ylim, expand = FALSE, crs = crs_robin, ndiscr = 1000) +
     cowplot::theme_minimal_grid() + 
     colorspace::scale_fill_continuous_sequential(
          palette = "Blues", rev = TRUE,
          na.value = "grey60", trans = "log10",
          breaks = leg_breaks,
          name = NULL,
          guide = guide_colorbar(frame.colour = "black",
                                 barwidth = 10,
                                 label.vjust = 2.5) # weirdly, moves labels closer to bar
     ) +
     theme(legend.direction = "horizontal",
           legend.position = c(.4, .2),
           plot.title = element_text(hjust = 0.5),
           plot.caption = element_text(face = "italic") ) +
     labs(title = paste("Number user visits", params$startdate, "thru", params$enddate),
          caption = "Countries with no visits in gray")
world_plot

ggsave(filename = here::here("plots", paste(params$enddate, "country_visits.png", sep = "_") ),
       plot = world_plot,
       width = 10,
       height = 6)
```

## Plot of use in United States

```{r us, echo = FALSE, fig.height = 6, fig.width = 10}
# US map
# Using urbnmapr, from GitHub, https://urbaninstitute.github.io/urbnmapr/
states_sf = urbnmapr::get_urbn_map("states", sf = TRUE)

# Add in user info to us map data
state_users = left_join(states_sf, us_states, by = c("state_name" = "region"))

# Not adjusting for population
state_plot = ggplot(data = state_users) + 
     geom_sf(aes(fill = users), color = "black", size = 0.5/.pt) +
     cowplot::theme_minimal_grid() + 
     colorspace::scale_fill_continuous_sequential(
          palette = "Blues", rev = TRUE,
          na.value = "grey60",
          breaks = c(10, 50, 100, 1000),
          name = NULL, trans = "sqrt",
          guide = guide_colorbar(frame.colour = "black",
                                 barwidth = 7.5,
                                 label.vjust = 2.5) # weirdly, moves labels closer to bar
     ) +
     theme(legend.direction = "horizontal",
           legend.position = c(.6, .05),
           plot.title = element_text(hjust = 0.5),
           plot.caption = element_text(face = "italic") ) +
     labs(title = paste("Number user visits", params$startdate, "thru", params$enddate),
          caption = "States with no visits in gray")
state_plot

ggsave(filename = here::here("plots", paste(params$enddate, "state_visits.png", sep = "_") ),
       plot = state_plot,
       width = 10,
       height = 6)
```

